Lessons Learned from the Fast Forward Storage and I/O Project

### 2016-07-18 paper.tex
added reviews and "plagarism" concerns (fixed via refs). Need to edit a bit more still.

\documentclass[conference]{IEEEtran}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

@@

\hyphenation{sub-trans-ac-tion}

@@

%\conferenceinfo{Cluster'14,} {September 17, 2014, Madrid, Spain}
%\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications}

\author{
\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}HDF Group koziol@hdfgroup.org}
\IEEEauthorblockA{\IEEEauthorrefmark{4}EMC john.bent@emc.com}
\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
}
\maketitle

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase 1 of the project complete, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examines the overall design.
Each layer of the proposed stack is discussed in some detail along with
cross-cutting topics that impact multiple layers, such as transactions and
metadata management. We also describe the proposed features to better support
some classes of big data applications.

This paper not only provides a timely summary of important aspects of the
design specifications but also captures the underlying reasoning that is not
available elsewhere. With this information, we encourage the broader community
to understand the design, intent, and future directions to foster discussion
guiding phase 2 and the ultimate production storage stack based on this work.
An initial performance evaluation of the early prototype implementation is also
provided to validate the presented design.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve extreme scale science platform
requirements.  Adding to the problem complexity is the variety of Big Data
problems users want to address using these platforms. Trying to address both of
these domains is a daunting problem.

To address these challenges, a joint effort between the US Department of
Energy's Office of Advanced Simulation and Computing and Advanced Scientific
Computing Research commissioned a project to develop a design and prototype for
an IO stack suitable for the extreme scale environment. It will be referred to
as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase recently completed with a second phase
currently getting underway as of early 2014. The core focus of the first phase
was basic functionality and design. While an idealized potential system would
be the perfect target architecture, the reality of budgets has tempered many of
the decisions. For example, extensive availability of NVRAN or SSDs on all of
the compute nodes is currently not economically feasible limiting some of the
potential design choices.  With this in mind, the second phase will refine this
design incorporating fault recovery and other features missing from the first
phase.

@@

\begin{figure}[htbp]
\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/arch-mapping}
\vspace{-0.15in}
\caption{Target Architecture and Component Mapping}
\label{fig:arch-mapping}
\vspace{-0.15in}
\end{figure}

At a more detailed view, the various layers of the IO stack each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is generally a high level IO library, such as the demonstration
HDF5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers and to enable advanced functionality that would
require more direct user intervention. This layer incorporates the necessary
features for ACGs from a end-user's perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack (in green) and offers features like Burst
Buffers to insulate the persistent storage array from bursty IO workloads.
Ideally, the IOD layer's functionality can be optional based on available
hardware and compute power provided on the IO Nodes (IONs). Much of the
functionality offered at this layer would shift either up or down the stack as
discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and translation layer between
the user-visible object model and the requirements of the underlying storage
infrastructure. It is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albeit with the potential performance penalty of manipulating
the shared, persistent storage array. At the bottom is the Versioning Object
Storage Device (VOSD) (in purple).  It serves as the interface for storing
objects of all types efficiently for each storage device in the parallel
storage array. Think of this layer as the physical disk interface layer. In
terms of Lustre, this would replace the API on individual storage devices with
an interface friendlier to the containers of objects and transactions/epochs
concepts used in the higher layers.

Along with analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product.  These ideas are presented to give a
fuller picture of where the project is going rather than dwelling on any
limitations of the published documents. This is most important to illustrate
how different concepts will work across layers since that information is spread
across multiple documents and may lack a cohesive overall view.

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the context of the HDF5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
layer functions. As with the IOD layer, the cross-cutting features will be
mentioned, but discussed more fully in the cross-cutting section. The VOSD
layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage. Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}. Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}. This shows that the prototype system based on the
proposed design can function. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

Many projects over the last couple of decades have sought to address some
challenging aspect of parallel file system design. The recent rise of ``Big
Data'' applications with different characteristic IO patterns have somewhat
complicated the picture. Extreme scale machines will be expected to handle both
the traditional simulation-related workloads as well as applications more
squarely in the Big Data arena. This will require some adjustments to the
underlying system for good performance for both scenarios.

@@

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems. Ceph's unique striping approach uses pseudo-random numbers with a
known seed eliminating the need for the metadata service to track where each
stripe in a parallel file is placed.

PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle. It has been
commercialized in recent years as OrangeFS.

Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.

GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM.

Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
striping system that detects the need for additional stripes for performance
and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly. The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file, but distributing the set of
files across a wide array of storage devices. The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.  These other file systems are mainly of
interest in the context of the ACG features of FFSIO and will be discussed more
in Section~\ref{sec:acg}.

@@

Since the proposal specifies a high-level IO API will be the primary end-user
interface for programmatically interacting with the FFSIO stack, the team used
the HDF5 API and leveraged the Virtual Object Layer (VOL) for the initial
design and implementation demonstration. This also serves as a good test
determining what are strictly necessary extensions to an existing IO API to
support the new functionality.  The additional functionality, such as
transactions, can be ignored for legacy implementations, but these applications
will not be able to take advantage of the asynchronous IO support inherent to
the new API.  The additions comprise (Figure~\ref{fig:vol-arch}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt

@@

\item\label{item:fs}
  Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
  the application developer with the capability of sending computation down to
  the IONs and get back results and perform other operations such as indexing
  and data reorganization for more efficient retrieval.

\item
  Analysis Shipping from CN to IONs or DAOS nodes. This is similar
  to~\ref{item:fs} but instead of returning the result over the network, it
  gets stored on the nodes and pointers to it are returned.

@@

HDF5~\cite{hdf5} has a versatile data model offering complex data objects
and metadata. Its information set is a collection of datasets, groups,
datatypes and metadata objects. The data model defines mechanisms for
creating associations between various information items. The main
components of HDF5 are described below.

@@

\item
  \textbf{Group}: A group is an object allowing association between HDF5
  objects. It is synonymous with directories in a file system. A group
  could contain multiple other groups, datasets, datatypes or attributes within
  it.

\item
  \textbf{Dataset}: HDF5 datasets are objects representing actual data
  or content. Datasets are arrays with potentially multiple dimensions. A
  dataset is characterized by a dataspace and a datatype. The dataspace
  captures the rank (number of dimensions) and the current and maximum
  extent in each dimension. The datatype describes the type of its data
  elements.

@@

\begin{figure}[htbp]
\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/vol-arch.png}
\vspace{-0.20in}
\caption{Architectural view of the VOL abstraction mechanism}
\label{fig:vol-arch}
\vspace{-0.10in}
\end{figure}

The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
running on the system's compute nodes. Using the VOL architecture, the IOD VOL
plugin uses a function shipper (RPC library) to forward the VOL calls to a
server component running on the IO nodes (IONs). This function shipping is the
IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
and executed at the IONs.

@@

Since HDF5 has traditionally offered an interface focused on files and the
internal data types, such as datasets, these concepts must be mapped onto the
proposed FFSIO data storage concepts. At a high level, think of the HDF5 types
mapping to FFSIO types as follows: file $\rightarrow$ container, dataset $\rightarrow$
array, and groups and attributes $\rightarrow$ key-value store.
%This mapping is shown in Table~\ref{tab:mapping}

%\begin{table}[ht]
%    \vspace{-0.10in}
%    \centering
%    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
%    \bigskip
%    \vspace{-0.15in}
%
%    \begin{tabular}{|r|r|}
%\hline
%HDF5 & FFSIO\\
%\hline
%file & container \\
%dataset & array \\
%group & key-value store \\
%attribute(s) & key-value store \\
%\hline
%    \end{tabular}
%    \label{tab:mapping}
%\end{table}
%
In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.

@@

The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
the massive process count on the storage stack. A current trend mixing MPI with
a threading library like OpenMP is addressing the same issue, but limited to
handling the parallelism on a single node rather than multiple nodes. Projected
extreme scale platforms will have far fewer storage stack end-points per
compute process in which to receive requests and data. By reducing the number
of simultaneous requests, delays can be reduced. This has been demonstrated for
the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
degree for accessing the storage devices
themselves~\cite{lofstead:2010:io-variability}.  The BlueGene platform
incorporated dedicated hardware to perform this role. The proposed
functionality for this layer, beyond managing the number of connections to the
IOD layer, is to implement function shipping from the compute nodes to the IO
nodes.

@@

\noindent\textbf{Design Philosophy}

>>>>

\documentclass[conference]{IEEEtran} \pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\usepackage{hyperref}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

@@

@@

%\title{Fast Forwarding Storage for Extreme Scale Platforms and Big Data Applications}
\title{Lessons Learned from the Fast Forward Storage and I/O Project}

\author{
%\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Lawrence Berkeley National Laboratory koziol@lbl.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Segate Government Solutions John.Bent@seagategov.com}
\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
}
\maketitle

\begin{abstract}
%The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
%project is going to have significant impact on storage systems design within
%and beyond the HPC community. With phase two of the project starting, it is an
%excellent opportunity to explore the complete design and how it will address
%the needs of extreme scale platforms.  This paper examines the overall design.
%Each layer of the proposed stack is discussed in some detail along with
%cross-cutting topics that impact multiple layers, such as transactions and
%metadata management. We also describe the proposed features to better support
%some classes of big data applications.
%
%This paper not only provides a timely summary of important aspects of the
%design specifications but also captures the underlying reasoning that is not
%available elsewhere. With this information, we encourage the broader community
%to understand the design, intent, and future directions to foster discussion
%guiding phase two and the ultimate production storage stack based on this work.
%An initial performance evaluation of the early prototype implementation is also
%provided to validate the presented design.

The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase two of the project starting, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examines each layer of the
proposed stack in some detail along with cross-cutting topics, such as
transactions and metadata management.

This paper not only provides a timely summary of important aspects of the
design specifications but also captures the underlying reasoning that is not
available elsewhere. We encourage the broader community to understand the
design, intent, and future directions to foster discussion guiding phase two
and the ultimate production storage stack based on this work.  An initial
performance evaluation of the early prototype implementation is also provided
to validate the presented design.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve extreme scale science platform
requirements.  While new hardware, such as non-volatile memory will help, we
still need a new software stack to incorporate this new hardware as well as
address the extreme parallelism and performance requirements demanded by
exascale applications. Adding to the problem complexity is the variety of Big
Data problems users want to address using these platforms. Unlike the
centralized storage arrays favored for HPC platforms, big data analytics
systems have grown up using storage distributed on all of the nodes driving a
very different software architecture. With post-exascale platforms required to
address both workloads, a new storage stack is required.

To address these challenges, a joint effort between the US Department of
Energy's Office of Advanced Simulation and Computing and Advanced Scientific
Computing Research commissioned a project to develop a design and prototype for
an IO stack suitable for the extreme scale environment. It will be referred to
as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase completed in 2014 with a second phase
currently getting underway. The core focus of the first phase was basic
functionality and design. While an idealized potential system would be the
perfect target architecture, the reality of budgets has tempered many of the
decisions. For example, extensive availability of NVRAM or SSDs on all of the
compute nodes is currently not economically feasible limiting some of the
potential design choices.  With this in mind, the second phase will refine this
design incorporating fault recovery and other features missing from the first
phase.

@@

\begin{figure}[htbp]
%\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/arch-mapping}
%\vspace{-0.15in}
\caption{Target Architecture and Component Mapping}
\label{fig:arch-mapping}
%\vspace{-0.15in}
\end{figure}

At a more detailed view, the various layers of the IO stack each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is comprised of generally a high level IO library, such as the
demonstration HDF5 library~\cite{hdf5} and a more complex API for accessing the
lower level components. This layer is in dark blue. Becuase the system supports
more complex architectures and supports richer functionality, hiding this
complexity behind a user-friendly API is the intent. It is possible to 
access the storage stack through the more complex API, but the additional
requirements beyond standard POSIX calls will prompt most users to use an IO
library.  This layer incorporates the necessary features for ACGs from a
end-user's perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. One special function
of note for this layer is that it is where function shipping will be deployed.
This is discussed in Section~\ref{sec:end-user}. The next two layers have
considerable functionality.

The IO Dispatcher (IOD) serves as the primary storage interface for the IO
stack (in green) and offers features like Burst Buffers to insulate the
persistent storage array from bursty IO workloads.  Ideally, the IOD layer's
functionality can be optional based on available hardware and compute power
provided on the IO Nodes (IONs). Transactions are handled primarily at this
layer. Much of the functionality offered at this layer would shift either up or
down the stack as discussed in detail below.

The Distributed Asynchronous Object Storage (DAOS) layer serves as the
persistent storage interface and translation layer between the user-visible
object model and the requirements of the underlying storage infrastructure. 
Transactions work a bit differently at this layer and are called epochs to
distinguish them. DAOS is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albeit with the potential performance and usability penalty of
manipulating the shared, persistent storage array directly.

At the bottom is the Versioning Object Storage Device (VOSD) (in purple).  It
serves as the interface for storing objects of all types efficiently for each
storage device in the parallel storage array. Think of this layer as the
physical disk interface layer. In terms of Lustre, this would replace the API
on individual storage devices with an interface friendlier to the containers of
objects and transactions/epochs concepts used in the higher layers.

Along with analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate system.  This information was gathered through
interviews with the core FFSIO team members. These ideas are presented to give
a fuller picture of where the project is going rather than dwelling on any
limitations of the published documents. This is most important to illustrate
how different concepts will work across layers since that information is spread
across multiple documents and may lack a cohesive overall view. Previously,
a poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
consistency and fault tolerance~\cite{lofstead:2014:ffsio-consistency} have
been presented.

The rest of the paper is organized as follows. An overview of related work is
presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the context of the HDF5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
layer functions. As with the IOD layer, the cross-cutting features will be
mentioned, but discussed more fully in the cross-cutting section. The VOSD
layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage. Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}. Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}. This shows that the prototype system based on the
proposed design can function. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

Many projects over the last couple of decades have sought to address some
challenging aspect of parallel file system design. The recent rise of ``Big
Data'' applications with different characteristic IO patterns have somewhat
complicated the picture. Vendors are shifting products to address the far
larger market forcing HPC systems to adapt to these different storage
approaches. Extreme scale machines will be expected to handle both the
traditional simulation-related workloads as well as applications more
squarely in the Big Data arena. This will require some adjustments to the
underlying system for good performance for both scenarios.

@@

Lustre~\cite{braam:lustre-arch} is the de facto standard on most major clusters
offering scalable performance and fine-grained end-user and programmatic
control over how data is placed in the storage system. The broad community
support has led to a solid code base with sufficient optimizations to serve as
the low-cost, proven solution. For each installation, system-wide settings that
apply to all files on the file system are made. The end user, should they have
different needs can reconfigure these characteristics on a file-by-file basis.
This becomes an issue because the dominant file size is tiny. In many cases, it
can be $<$ 4 KB.  To keep from slowing the overall system performance when
creating and opening these files, most systems are configured to use a 1 MB
stripe size and a stripe count of 4 meaning only 4 storage targets are used per
file. This limits the default aggregate bandwidth to the combined speed of four
storage targets. By reconfiguring on a file-by-file basis, this default can be
overcome for large, parallel files achieving very high performance. The
downside is that this setting must be done to take advantage of the full
parallel file system performance.

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in
parallel file systems. Ceph's unique striping approach uses pseudo-random
numbers with a known seed eliminating the need for the metadata service to
track where each stripe in a parallel file is placed. Ceph's strengths are in
providing good perforamnce and scalability with the ability to handle failures
and deploying new storage adapting the system in a live environment. However,
this failure handling advantage was shown~\cite{wang:2013:ceph} to limit peak
performance more than other systems like Lustre.  More recently, some of these
limitations have begun to be addressed by the Ceph team, but no new evaluations
have been performed to determine if these changes close the gap sufficiently to
address the extreme HPC performance needs.

PVFS~\cite{carns:pvfs} was built understanding the scalibility bottlenecks
Lustre suffers. For example, Lustre requires all processes opening a file to
hit the metadata server to receive a proper file handle. PVFS reduces this load
by allowing a single process to open a file and sharing the handle with other
processes participating in the IO operation. There are also other optimizations
that enhance file system performance. It has been commercialized in recent
years as OrangeFS.

GPFS~\cite{schmuck:gpfs} offers a highly scalable parallel file system with
robust functionality to handle both parallel storage, recovery, and
optimization. It only supports a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM. Unfortunately, the stripe size is fixed introducing potentially false
sharing, when two processes indpendently write to the same stripe, but without
overlapping, causing potentially reduced performance. Beyond these sorts of
fixed parameters, a wide variety of optimizations and features are available
for additional licensing fees.

Panasas~\cite{panasas:architecture} uses a fundamentally different approach to
parallel file system performance. When parallel writers simultaneously write to
a shared file, the system dynamically adapts the number of stripes to maintain
high performance. This adaption is invisible to the user other than seeing that
the system maintains high performance no matter what configuration the workload
exhibits.

This project learns from all of these parallel file system efforts and offers a
scalable approach that can work well for everything from a small cluster to the
largest exascale platforms. By understanding what works well and what the
limitations are for each of the above systems as well as emerging hardware
architectures, this project addresses the limitations while maintaining the
advantages of the above systems.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly. The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file typically on a single storage
device. For performance, the file or object may be replicated.  The other,
popular distributed file system of note is NFS~\cite{powlowski:1994:nfs3} that
has been used for decades for enterprise file systems.  NFS is known to support
a global namespace with data migrating towards users on access and pushed
towards safer storage based on local platform characteristics. These other file
systems are mainly of interest in the context of the ACG features of FFSIO and
will be discussed more in Section~\ref{sec:acg}.

The main alternative from scratch design for a file system is
Sirocco~\cite{sirocco,curry:2016:sirocco}. Rather than continuing the striped
design of existing parallel file systems, Sirocco is inspired by peer-to-peer
and object-based systems and includes features like transactions to protect
data modification process independence when writing to avoid coordination
overhead.  The base assumptions are that storage is pervasive and volatile.
Storage devices and locations may come and go randomly, reminiscent of the
Google or Ceph assumptions of regularly failing hardware. When data is pushed
into the system, initial resilience characteristics are guaranteed prior to
returning control back to the user. Then, as system pressures dictate, data
will either replicate as demanded by use and/or migrated towards long-term
resilience requirements.  Unlike the FFSIO project, Sirocco assumes it is
possible that data may be successfully stored in the system, but it is
currently inaccessible because all copies are currently offline. There is also
some potential difficulty in finding data since it will migrate around the
system. To be fair, Sirocco intends to function as the storage layer for a
higher level file system API making many of the awkward system features
invisible to the end user. Sirocco is in the process of being released
publicly.

@@

Since the proposal specifies a high-level IO API will be the primary end-user
interface for programmatically interacting with the FFSIO stack, the team used
the HDF5 API and leveraged its Virtual Object Layer (VOL) for the initial
design and implementation demonstration. This also serves as a good test
determining what are strictly necessary extensions to an existing IO API to
support the new functionality.  The additional functionality, such as
transactions, can be ignored for legacy implementations, but these applications
will not be able to take advantage of the asynchronous IO support inherent to
the new API.  The additions comprise (Figure~\ref{fig:vol-arch}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
%\itemsep1pt\parskip0pt\parsep0pt

@@

\item
  Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
  the application developer with the capability of sending computation down to
  the IONs and get back results and perform other operations such as indexing
  and data reorganization for more efficient retrieval.

\item
  Analysis Shipping from compute nodes to IO Nodes or DAOS nodes. This is
  similar to function shipping, but instead of returning the result over the
  network, it is stored on the nodes and pointers to the data are returned.

@@

HDF5~\cite{hdf5} has a versatile data model offering complex data objects
and metadata. Its information set is a collection of datasets, groups,
datatypes and metadata objects. The data model defines mechanisms for
creating associations between various information items. The main
conceptual components for data stored in HDF5 are described below.

@@

\item
  \textbf{Group}: A group is an object allowing association between HDF5
  objects. It is synonymous with directories in a file system. A group
  could contain multiple other groups, datasets, datatypes or attributes within
  it. Groups are named and then accessed using a standard path notation
  similar to Linux with a ``/'' separating each group name in the hierarchy
  from the root to the nested group of interest.

\item
  \textbf{Dataset}: HDF5 datasets are objects representing actual data
  or content. Datasets are typically arrays with potentially multiple
  dimensions. Other types, such as strings and scalars, are also possible. A
  dataset is characterized by a dataspace and a datatype. The dataspace
  captures the rank (number of dimensions) and the current and maximum
  extent in each dimension. The datatype describes the type of its data
  elements. By default, the entire data set is stored as a single chunk
  reassembled from all processes. It is possible to use a uniform chunking
  format where data is stored in fixed sized chunks instead.

@@

\begin{figure}[htbp]
%\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/vol-arch.png}
%\vspace{-0.20in}
\caption{Architectural view of the VOL abstraction mechanism}
\label{fig:vol-arch}
%\vspace{-0.10in}
\end{figure}

The IOD VOL plugin serves as the bridge between HDF5 and the IOD Layer
(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
running on the system's compute nodes. Using the VOL architecture, the IOD VOL
plugin uses a function shipper (RPC library) to forward the VOL calls to a
server component running on the IO nodes (IONs). This function shipping is the
IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
and executed at the IONs.

Since the IOD layer is optional by design, a second VOL plugin is required to
access DAOS directly. This additional complexity was deemed acceptable for the
flexibility it affords. Further, since all end user interactions are intended
to be through an IO API, having two different plugins is a small amount of
extra work to provide a single interface that would operate on widely different
scale deployments (e.g., one too small to have the IOD layer and one at the
other extreme with a large IOD layer interfacing with a large, shared DAOS
layer).

@@

Since HDF5 has traditionally offered an interface focused on files and the
internal data types, such as datasets, these concepts must be mapped onto the
proposed FFSIO data storage concepts.  This mapping is shown in
Table~\ref{tab:mapping}

\begin{table}[ht]
%    \vspace{-0.10in}
    \centering
    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
    \bigskip
%    \vspace{-0.15in}

    \begin{tabular}{|r|r|}
\hline
HDF5 & FFSIO\\
\hline
file & container \\
dataset & array \\
group & key-value store \\
attribute(s) & key-value store \\
\hline
    \end{tabular}
    \label{tab:mapping}
\end{table}

In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.

@@

The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
the massive process count fan in onto the storage stack. The current trend of
using both MPI and a node-level threading library like OpenMP, CUDA, or OpenACC
is addressing the same issue, but limited to handling the parallelism on a
single node rather than multiple nodes. Projected extreme scale platforms will
have far fewer storage stack end-points per compute process or even compute
node in which to receive requests and data. By reducing the number of
simultaneous requests, delays can be reduced. This has been demonstrated for
the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
degree for accessing the storage devices
themselves~\cite{lofstead:2010:io-variability}.  The BlueGene platform
incorporated dedicated hardware to perform this role. The proposed
functionality for this layer, beyond managing the number of connections to the
IOD layer, is to implement function shipping from the compute nodes to the IO
nodes.

@@

\subsection{Data Versioning}
